---
title: ""
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{patentsview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# patentsview

> An R Client to the PatentsView API

## Installation

```{r eval = FALSE}
devtools::install_github("ropensci/patentsview")
```

## Basic usage

The [PatentsView API](http://www.patentsview.org/api/doc.html) provides an interface to a disambiguated version of USPTO. The `patentsview` R package provides one main function, `search_pv()`, to make it easy to interact with that API:

```{r}
library(patentsview)

search_pv(query = '{"_gte":{"patent_date":"2007-01-01"}}',
          endpoint = "patents")
```

## Learning more

Check out the:

* [Getting started vignette](articles/getting-started.html) if you are a first-time user
* In-depth tutorial on [writing queries](articles/writing-queries.html)
* List of [examples](articles/examples.html)
* WIP: Applications of the data (e.g., finding [top assignees](articles/assignees.html))