---
title: "patentsview"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<br>

## Installation

You can get the stable version from CRAN:

```{r eval = FALSE}
install.packages("patentsview")
```

Or the development version from GitHub:

```{r eval = FALSE}
if (!"devtools" %in% rownames(installed.packages())) 
  install.packages("devtools")

devtools::install_github("ropensci/patentsview")
```

## Basic usage

The [PatentsView API](http://www.patentsview.org/api/doc.html) provides an interface to a disambiguated version of USPTO. The `patentsview` R package provides one main function, `search_pv()`, to make it easy to interact with the API:

```{r}
library(patentsview)

search_pv(query = '{"_gte":{"patent_date":"2007-01-01"}}')
```

## Learning more

Check out the:

* [Getting started vignette](articles/getting-started.html) or [rOpenSci blog post](https://ropensci.org/blog/blog/2017/09/19/patentsview) if you are a first-time user
* In-depth tutorial on [writing queries](articles/writing-queries.html)
* List of [basic examples](articles/examples.html)
* Examples of data applications (e.g., a brief analysis of the [top assignees](articles/top-assignees.html) in the field of databases)
