---
output: github_document
title: "patentsview"
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

> An R Client to the PatentsView API

[![](http://badges.ropensci.org/112_status.svg)](https://github.com/ropensci/onboarding/issues/112)
[![Linux Build Status](https://travis-ci.org/ropensci/patentsview.svg?branch=master)](https://travis-ci.org/ropensci/patentsview)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ropensci/patentsview?branch=master&svg=true)](https://ci.appveyor.com/project/ropensci/patentsview)

## Installation

You can get the stable version from CRAN:

```{r eval = FALSE}
install.packages("patentsview")
```

Or the development version from GitHub:

```{r eval = FALSE}
if (!("devtools" %in% installed.packages()[,1]))
  install.packages("devtools")

devtools::install_github("ropensci/patentsview")
```

## Basic usage

The [PatentsView API](http://www.patentsview.org/api/doc.html) provides an interface to a disambiguated version of USPTO. The `patentsview` R package provides one main function, `search_pv()`, to make it easy to interact with that API:

```{r}
library(patentsview)

search_pv(query = '{"_gte":{"patent_date":"2007-01-01"}}',
          endpoint = "patents")
```

## Learning more

Check out the:

* [Getting started vignette](articles/getting-started.html) if you are a first-time user
* In-depth tutorial on [writing queries](articles/writing-queries.html)
* List of [examples](articles/examples.html)
* WIP: Data appliations (e.g., finding [top assignees](articles/assignees.html))