---
title: "patentsview"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

> An R client to the PatentsView API

[![](http://badges.ropensci.org/112_status.svg)](https://github.com/ropensci/software-review/issues/112)
[![R-CMD-check](https://github.com/ropensci/patentsview/workflows/R-CMD-check/badge.svg)](https://github.com/ropensci/patentsview/actions)
[![CRAN version](http://www.r-pkg.org/badges/version/patentsview)](https://cran.r-project.org/package=patentsview)
[![runiverse-package](https://ropensci.r-universe.dev/badges/patentsview)](https://ropensci.r-universe.dev/patentsview)

## Installation

You can get the stable version from CRAN:

```{r eval = FALSE}
install.packages("patentsview")
```

### Development version

To get a bug fix or to use a feature from the development version, you can install the development version from GitHub:

```{r eval = FALSE}
# install.packages("pak")
pak::pak("ropensci/patentsview")
```

## Basic usage

The [PatentsView API](<https://search.patentsview.org/docs/docs/Search%20API/SearchAPIReference/>) provides an interface to a disambiguated version of USPTO. The `patentsview` R package provides one main function, `search_pv()`, to make it easy to interact with the API:

```{r}
library(patentsview)

search_pv(query = '{"_gte":{"patent_date":"2007-01-01"}}')
```

## Learning more

Head over to the package's [webpage](https://docs.ropensci.org/patentsview/index.html) for more info, including:

* A [getting started vignette](https://docs.ropensci.org/patentsview/articles/getting-started.html) for first-time users. The package was also introduced in an [rOpenSci blog post](https://docs.ropensci.org/patentsview/articles/ropensci-blog-post.html), rewritten for the new version of the API.
* An in-depth tutorial on [writing queries](https://docs.ropensci.org/patentsview/articles/writing-queries.html)
* A list of [basic examples](https://docs.ropensci.org/patentsview/articles/examples.html)
* Two examples of data applications (e.g., a brief analysis of the [top assignees](https://docs.ropensci.org/patentsview/articles/top-assignees.html) in the field of databases)
