---
title: "Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{patentsview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Which patents have been cited by more than 500 US patents?

```{r eval = params$eval_all}
library(patentsview)

search_pv(query = qry_funs$gt(patent_num_cited_by_us_patents = 500))
```

How many distinct inventors (disambiguated) are represented by these highly-cited patents?

```{r eval = params$eval_all}
# Setting subent_cnts = TRUE will give us the subentity counts. Since inventors 
# are subentities for the patents endpoint, this means we will get their counts.
search_pv(query = qry_funs$gt(patent_num_cited_by_us_patents = 500),
          fields = c("patent_number", "inventor_id"), subent_cnts = TRUE)
```

What patents has Microsoft (disambiguated) published since 2010?

```{r eval = params$eval_all}
query <- with_qfuns(
  and(
    gte(patent_date = "2010-01-01"),
    contains(assignee_organization = "microsoft")
  )
)
search_pv(query = query)
```

Which assignees have an inventor whose last name contains "smith" (e.g., "smith", "johnson-smith")? Also, give me the patent data where those "smiths" occur.

```{r eval = params$eval_all}
# Get all possible retrievable assignee-level and patent-level fields for 
# the assignees endpoint:
asgn_pat_flds <- get_fields("assignees", c("assignees", "patents"))

# Ask the PatentsView API for these fields:
search_pv(query = qry_funs$contains(inventor_last_name = "smith"), 
          endpoint = "assignees", fields = asgn_pat_flds)
```

What are the top ten CPC subsections for patents funded by the DOE?

```{r eval = params$eval_all}
search_pv(query = qry_funs$contains(govint_org_name = "department of energy"), 
          endpoint = "cpc_subsections", 
          fields =  "cpc_total_num_patents",
          sort = c("cpc_total_num_patents" = "desc"), 
          per_page = 10)
```