---
title: "Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{patentsview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Patents endpoint

Which patents have been cited by more than 500 US patents?

```{r}
library(patentsview)

search_pv(query = qry_funs$gt(patent_num_cited_by_us_patents = 500))
```

How many distinct inventors are represented by these highly-cited patents?

```{r}
# Setting subent_cnts = TRUE will give us the subentity counts. Since inventors 
# are subentities for the patents endpoint, this means we will get their counts.
search_pv(query = qry_funs$gt(patent_num_cited_by_us_patents = 500),
          fields = c("patent_number", "inventor_id"), subent_cnts = TRUE)
```

Where geographically have Microsoft inventors been coming from over the past 20 years?

```{r}
# Write the query
query <- with_qfuns(
  and(
    gte(patent_date = "2007-07-25"), # Dates are in yyyy-mm-dd format 
    contains(assignee_organization = "microsoft")
  )
)

# Create a field list
inv_fields <- get_fields(endpoint = "patents", groups = "inventors")
fields <- c(inv_fields, "patent_number")

# Pull the data
pv_out <- search_pv(query = query, fields = fields, all_pages = TRUE)

# Unnest the inventor list column
unnest_pv_data(data = pv_out$data, pk = "patent_number")
```

## Inventors endpoint

Return patent numbers and titles with at least one inventor whose location was Chicago, IL.<sup><a href="#fn1" id="ref1">1</a></sup>

```{r}
search_pv(query = '{"_and":[{"location_city":"Chicago"},{"location_state":"IL"}]}',
          endpoint = "inventors", fields = c("patent_number", "patent_title"))
```

## Assignees endpoint

Which assignees have an interest in beer?

```{r}
search_pv(query = qry_funs$text_phrase(patent_title = "beer"), 
          endpoint = "assignees")
```

*** 
<sup id="fn1">1</sup> Example taken from http://www.patentsview.org/api/inventor.html<sup><a href="#ref1">back</a></sup>