---
title: "Citation networks"
output: rmarkdown::html_vignette
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, 
  comment = "#>", 
  warning = FALSE,
  message = FALSE
)
```

The following is a brief foray into patent citation networks. The analysis is done on 3 patents that describe patent citation analysis (PCA) themselves. 

The first step is to download the relevant data from the PatentsView API. We can use the CPC code of  [Y10S707/933](https://worldwide.espacenet.com/classification#!/CPC=Y10S707/933) to pull PCA patents. 


```{r}
library(patentsview)
library(dplyr)
library(visNetwork)
library(magrittr)
library(stringr)

# Write a query to get the patents that are assigned a CPC code of "Y10S707/933": 
query <- qry_funs$begins(cpc_subgroup_id = "Y10S707/933")

# Create a list of fields to pull from the API:
fields <- c(
  "patent_number", 
  "patent_title",
  "cited_patent_number", # Which patents do they cite?
  "citedby_patent_number" # Which patents cite them?
)

# Send a request to the API:
res <- search_pv(query = query, fields = fields, all_pages = TRUE)

# Unnest the data found in the list columns:
res_lst <- unnest_pv_data(res$data, pk = "patent_number")
res_lst
```

There are only `r nrow(res_lst$patents)` PCA patents. These patents cite `r nrow(res_lst$cited_patents)` patents and are cited by `r nrow(res_lst$citedby_patents)` patents. Let's visualize the citations among the PCA patents. We'll create our visualization using the `visNetwork` package, which requires us to create a data frame of nodes and a data frame of edges.

```{r}
pat_title <- function(title, number) {
  temp_title <- str_wrap(title)
  i <- gsub("\\n", "<br>", temp_title)
  paste0('<a href="https://patents.google.com/patent/US', number, '">', i, '</a>')
}

edges <-
  res_lst$cited_patents %>%
    semi_join(x = ., y = ., by = c("cited_patent_number" = "patent_number")) %>%
    set_colnames(c("from", "to"))

nodes <-
  res_lst$patents %>%
    mutate(
      id = patent_number,
      label = patent_number,
      title = pat_title(patent_title, patent_number)
    )

visNetwork(nodes = nodes, edges = edges, height = "400px", width = "100%",
           main = "Citations among patent citation analysis (PCA) patents") %>%
  visEdges(arrows = list(to = list(enabled = TRUE))) %>%
  visIgraphLayout()
```

It looks like several of the patents cite patent number 6,499,029, perhaps indicating that patent 6,499,029 describes technology that is foundational to this small field. However, when we hover over the nodes we see that several of the patents have the same title. Clicking on the titles brings us to their full text on Google Patents, which confirms that several of the PCA patents belong to the same patent family.[^1]  Let's choose one of the patents in each family to act as the family's representative. This will reduce the size of the subsequent network while hopefully retaining its overall structure.

```{r}
p3 <- c("7797336", "9075849", "6499026")
res_lst2 <- lapply(
  res_lst, function(x) x[x$patent_number %in% p3, ]
)
```

With only 3 patents, it will probably be possible to visualize how these patents' cited and citing patents are all related to one another. Let's create a list of these "relevant patents" (i.e., the 3 patents plus all of their cited and citing patents), and then get a list of all of _their_ cited patents. This list of cited patents will allow us to measure how similar the relevant patents are to one another. 

```{r}
rel_pats <-
  res_lst2$cited_patents %>%
    rbind(setNames(res_lst2$citedby_patents, names(.))) %>% 
    select(-patent_number) %>%
    rename(patent_number = cited_patent_number) %>%
    bind_rows(data.frame(patent_number = p3)) %>% 
    distinct() %>%
    filter(!is.na(patent_number))

# Look up which patents the relevant patents cite:
rel_pats_res <- search_pv(
  query = list(patent_number = rel_pats$patent_number),
  fields =  c("cited_patent_number", "patent_number", "patent_title"), 
  all_pages = TRUE,
  method = "POST"
)

rel_pats_lst <- unnest_pv_data(rel_pats_res$data, pk = "patent_number")
```

Now we know which patents the `r nrow(rel_pats_lst$patents)` relevant patents cite. This allows us to measure the similarity between any two relevant patents by seeing how many cited references they have in common (a method known as [bibliographic coupling](https://en.wikipedia.org/wiki/Bibliographic_coupling)).

```{r}
cited_pats <-
  rel_pats_lst$cited_patents %>%
    filter(!is.na(cited_patent_number))

full_network <- 
  cited_pats %>%
    do({
      .$ind <- group_by(., patent_number) %>%  
               group_indices()
      group_by(., patent_number) %>%  
              mutate(sqrt_num_cited = sqrt(n()))
    }) %>%
    inner_join(x = ., y = ., by = "cited_patent_number") %>%
    filter(ind.x > ind.y) %>%
    group_by(patent_number.x, patent_number.y) %>% 
    mutate(cosine_sim = n() / (sqrt_num_cited.x * sqrt_num_cited.y)) %>% 
    ungroup() %>%
    select(matches("patent_number\\.|cosine_sim")) %>%
    distinct()
```

`full_network` is an edge list where the similarity between two "relevant patents" is stored in the `cosine_sim` column. This edge list contains pairs of patents that share at least one cited reference in common, so it probably contains a lot of patent pairs that really aren't all that similar. Let's try to identify a natural level of `cosine_sim` to filter on, so our subsequent network plot is not enormous.

```{r, fig.height = 4, fig.width = 7}
hist(full_network$cosine_sim, main = "Similarity scores between patents relevant to PCA",
     xlab = "Cosine similarity", ylab = "Number of patent pairs")
```

There appears to be a smallish group of patent pairs that are very similar to each other (`cosine_sim` > 0.8), which makes it tempting to choose 0.8 as a cutoff point. However, patent pairs that have reference lists that are this similar to each other are probably just patents in the same family. Let's choose 0.1 as a cutoff point instead, as there doesn't appear to be all that many pairs above this point.[^2] 

```{r}
edges <- 
  full_network %>%
    filter(cosine_sim >= .1) %>% 
    rename(from = patent_number.x, to = patent_number.y, value = cosine_sim) %>%
    mutate(title = paste("Cosine similarity =", as.character(round(value, 3))))

nodes <-
  rel_pats_lst$patents %>%
    rename(id = patent_number) %>%
    mutate(
      color = ifelse(id %in% p3, "#97C2FC", "#DDCC77"),
      label = id,
      title = pat_title(patent_title, id)
    )

visNetwork(nodes = nodes, edges = edges, height = "700px", width = "100%",
           main = "Network of patents relevant to PCA") %>%
  visEdges(color = list(color = "#343434")) %>%
  visOptions(highlightNearest = list(enabled = TRUE, degree = 1)) %>%
  visIgraphLayout()
```

[^1]: A patent family is a group of related patents, usually all authored by the same inventor and relating to the same technology.
[^2]: This is still a pretty arbitrary choice. Take a look at algorithms like the [disparity filter](http://www.pnas.org/content/106/16/6483.full.pdf) for a more systematic way to filter edges.